---
title: "HER2 Integration Project - Data Quality Report"
subtitle: "Session 1: TCGA-BRCA Data Acquisition"
author: "Research Team"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 10
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(ggplot2)
library(plotly)
library(DT)
library(jsonlite)
library(knitr)
library(VennDiagram)
library(gridExtra)

# Load session metadata
metadata <- fromJSON("metadata/session1_metadata.json")
```

# Executive Summary

## Session Overview
- **Objective**: `r metadata$objective`
- **Date**: `r metadata$date`
- **Duration**: `r round(metadata$duration_minutes, 1)` minutes
- **Status**: ✅ Complete

## Key Metrics
```{r summary_metrics, echo=FALSE}
metrics_df <- data.frame(
  Metric = c(
    "Total Unique Patients",
    "Complete Data Cases", 
    "Clinical Data Available",
    "RNA Expression Available",
    "Copy Number Available",
    "Complete Data Percentage"
  ),
  Value = c(
    ifelse(is.null(metadata$quality_metrics$total_unique_patients), "TBD", metadata$quality_metrics$total_unique_patients),
    ifelse(is.null(metadata$sample_counts$complete_data), "TBD", metadata$sample_counts$complete_data),
    ifelse(is.null(metadata$sample_counts$total_clinical), "TBD", metadata$sample_counts$total_clinical),
    ifelse(is.null(metadata$sample_counts$total_rna), "TBD", metadata$sample_counts$total_rna),
    ifelse(is.null(metadata$sample_counts$total_cnv), "TBD", metadata$sample_counts$total_cnv),
    ifelse(is.null(metadata$quality_metrics$complete_data_percent), "TBD%", paste0(metadata$quality_metrics$complete_data_percent, "%"))
  )
)

kable(metrics_df, caption = "Data Acquisition Summary")
```

# Data Acquisition Details

## Datasets Created
```{r datasets_table, echo=FALSE}
if(length(metadata$datasets_created) > 0) {
  datasets_df <- data.frame(
    Dataset = metadata$datasets_created,
    Description = c(
      "Clinical data with patient characteristics and outcomes",
      "ERBB2 RNA expression levels from RNA-seq",
      "ERBB2 copy number alteration scores"
    )[1:length(metadata$datasets_created)],
    Status = "✅ Complete"
  )
  
  kable(datasets_df, caption = "Created Datasets")
} else {
  cat("No datasets created in this session.")
}
```

# Platform Data Analysis

```{r load_data, include=FALSE}
# Try to load the actual data files for analysis
clinical_data <- tryCatch(readRDS("data/processed/tcga_clinical.rds"), error = function(e) NULL)
rna_data <- tryCatch(readRDS("data/processed/tcga_expression.rds"), error = function(e) NULL)
cnv_data <- tryCatch(readRDS("data/processed/tcga_cnv.rds"), error = function(e) NULL)
```

## Clinical Data Overview
```{r clinical_overview}
if(!is.null(clinical_data)) {
  cat("Clinical data loaded successfully with", nrow(clinical_data), "patients\n")
  
  # Age distribution
  if("age_years" %in% names(clinical_data)) {
    age_plot <- ggplot(clinical_data, aes(x = age_years)) +
      geom_histogram(bins = 30, fill = "steelblue", alpha = 0.7) +
      labs(title = "Age Distribution", x = "Age (years)", y = "Count") +
      theme_minimal()
    
    print(age_plot)
  }
  
  # Sample characteristics table
  char_summary <- clinical_data %>%
    summarise(
      n_patients = n(),
      median_age = median(age_years, na.rm = TRUE),
      age_iqr_low = quantile(age_years, 0.25, na.rm = TRUE),
      age_iqr_high = quantile(age_years, 0.75, na.rm = TRUE),
      missing_age = sum(is.na(age_years)),
      male_count = sum(gender == "male", na.rm = TRUE),
      female_count = sum(gender == "female", na.rm = TRUE)
    )
  
  kable(char_summary, caption = "Clinical Data Summary", digits = 1)
} else {
  cat("Clinical data not available for analysis.")
}
```

## RNA Expression Analysis
```{r rna_analysis}
if(!is.null(rna_data)) {
  cat("RNA expression data loaded with", nrow(rna_data), "samples\n")
  
  # ERBB2 expression distribution
  expr_plot <- ggplot(rna_data, aes(x = erbb2_log2)) +
    geom_histogram(bins = 50, fill = "darkgreen", alpha = 0.7) +
    labs(title = "ERBB2 Expression Distribution", 
         x = "Log2(ERBB2 + 1)", y = "Count") +
    theme_minimal()
  
  print(expr_plot)
  
  # Expression summary statistics
  expr_summary <- rna_data %>%
    summarise(
      n_samples = n(),
      median_expression = median(erbb2_log2, na.rm = TRUE),
      q25_expression = quantile(erbb2_log2, 0.25, na.rm = TRUE),
      q75_expression = quantile(erbb2_log2, 0.75, na.rm = TRUE),
      min_expression = min(erbb2_log2, na.rm = TRUE),
      max_expression = max(erbb2_log2, na.rm = TRUE),
      zero_counts = sum(erbb2_counts == 0, na.rm = TRUE)
    )
  
  kable(expr_summary, caption = "ERBB2 Expression Summary", digits = 2)
} else {
  cat("RNA expression data not available for analysis.")
}
```

## Copy Number Analysis  
```{r cnv_analysis}
if(!is.null(cnv_data)) {
  cat("Copy number data loaded with", nrow(cnv_data), "samples\n")
  
  # CNV score distribution
  cnv_plot <- ggplot(cnv_data, aes(x = erbb2_cnv_score)) +
    geom_histogram(bins = 50, fill = "darkorange", alpha = 0.7) +
    labs(title = "ERBB2 Copy Number Distribution", 
         x = "Copy Number Score", y = "Count") +
    theme_minimal()
  
  print(cnv_plot)
  
  # CNV summary statistics
  cnv_summary <- cnv_data %>%
    summarise(
      n_samples = n(),
      median_cnv = median(erbb2_cnv_score, na.rm = TRUE),
      q25_cnv = quantile(erbb2_cnv_score, 0.25, na.rm = TRUE),
      q75_cnv = quantile(erbb2_cnv_score, 0.75, na.rm = TRUE),
      min_cnv = min(erbb2_cnv_score, na.rm = TRUE),
      max_cnv = max(erbb2_cnv_score, na.rm = TRUE),
      amplified_samples = sum(erbb2_cnv_score > log2(1.5), na.rm = TRUE)
    )
  
  kable(cnv_summary, caption = "ERBB2 Copy Number Summary", digits = 3)
} else {
  cat("Copy number data not available for analysis.")
}
```

# Data Overlap Analysis

```{r overlap_analysis}
if(!is.null(clinical_data) && !is.null(rna_data) && !is.null(cnv_data)) {
  
  # Create overlap summary
  clinical_ids <- clinical_data$patient_id
  rna_ids <- rna_data$patient_id  
  cnv_ids <- cnv_data$patient_id
  
  all_ids <- unique(c(clinical_ids, rna_ids, cnv_ids))
  
  overlap_summary <- data.frame(
    Platform = c("Clinical", "RNA Expression", "Copy Number"),
    Available_Samples = c(length(clinical_ids), length(rna_ids), length(cnv_ids)),
    Unique_Patients = c(length(unique(clinical_ids)), length(unique(rna_ids)), length(unique(cnv_ids)))
  )
  
  kable(overlap_summary, caption = "Platform Data Availability")
  
  # Overlap counts
  overlap_counts <- data.frame(
    Data_Combination = c(
      "Clinical Only",
      "Clinical + RNA", 
      "Clinical + CNV",
      "RNA + CNV",
      "All Three Platforms",
      "Total Unique"
    ),
    Patient_Count = c(
      length(setdiff(clinical_ids, union(rna_ids, cnv_ids))),
      length(intersect(clinical_ids, rna_ids)),
      length(intersect(clinical_ids, cnv_ids)), 
      length(intersect(rna_ids, cnv_ids)),
      length(intersect(intersect(clinical_ids, rna_ids), cnv_ids)),
      length(all_ids)
    )
  )
  
  kable(overlap_counts, caption = "Data Platform Overlaps")
  
} else {
  cat("Cannot perform overlap analysis - missing data files.")
}
```

# Quality Control Flags

## Missing Data Assessment
```{r missing_data}
missing_summary <- data.frame(
  Issue = c(
    "HER2 IHC Status Missing", 
    "Clinical Covariates Missing",
    "RNA Expression Outliers",
    "Copy Number Outliers",
    "Survival Data Missing"
  ),
  Status = c(
    "⚠️ To be extracted in Session 2",
    "⚠️ To be processed in Session 2", 
    "✅ Will be assessed",
    "✅ Will be assessed",
    "⚠️ To be validated"
  ),
  Action_Required = c(
    "Extract from clinical supplement files",
    "Harmonize clinical variables",
    "Apply statistical outlier detection", 
    "Apply statistical outlier detection",
    "Validate survival endpoints"
  )
)

kable(missing_summary, caption = "Quality Control Action Items")
```

## Next Session Preparation

### Files Ready for Session 2:
```{r next_session}
if(length(metadata$next_session_inputs) > 0) {
  cat("✅ Ready for Session 2:\n")
  for(file in metadata$next_session_inputs) {
    cat("  -", file, "\n")
  }
} else {
  cat("⚠️ No files ready - review Session 1 execution")
}
```

### Session 2 Objectives:
1. **Data Harmonization**: Standardize HER2 IHC categories and extract from clinical supplements
2. **Quality Control**: Implement comprehensive QC procedures and outlier detection  
3. **Missing Data**: Develop multiple imputation strategy for incomplete cases
4. **Clinical Variables**: Process ER/PR status, tumor grade, stage, and molecular subtypes
5. **Analysis Dataset**: Create final harmonized dataset ready for modeling

---

**Report Generated**: `r Sys.time()`  
**Next Session**: Data Harmonization & Quality Control